---
title: "Data Exploration"
author: "Su Goh"
date: "12/12/2020"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(lme4)
library(lmtest)
library(lmerTest)
```

```{r include=FALSE}
df <- st_read('data/CA16_CTs_all.gpkg')
df$percent_hh_with_children <- 100 * df$hh_with_children / df$hh_total
df$percent_drivers_female <- 100 * df$commute_driver_female / (df$commute_driver_male + df$commute_driver_female)
df$percent_publictransit_female <- 100 * df$commute_publictransit_female / (df$commute_publictransit_male + df$commute_publictransit_female)
df$lfp_gap <- df$lfp_male - df$lfp_female
df <- na.omit(df)
df_no_geom <- st_drop_geometry(df)
mtl_data <- filter(df, csd_uid == '2466023')
to_data <- filter(df, csd_uid == '3520005')
```

## Data Descriptives: DV
```{r desc1}
lfp <- rbind(data.frame(lfp = df$lfp_female, gender ='F'), data.frame(lfp = df$lfp_male, gender ='M'))
avg_lfp <- mean(df$lfp_female)
avg_lfp_m <- mean(df$lfp_male)
ggplot(lfp, aes(x=lfp, fill=gender)) + geom_histogram(alpha=0.5, position="identity") + geom_vline(xintercept=avg_lfp_m) + geom_vline(xintercept=avg_lfp)+  scale_x_continuous(breaks=seq(0,100,10)) + labs(x='LFP (%)') 
avg_gap <- mean(df$lfp_gap)
ggplot(df_no_geom, aes(lfp_gap)) + geom_histogram(color="black", fill="white", binwidth = 2) + geom_vline(xintercept = avg_gap, color='red') + labs(x='percentage points')
```


## Data Descriptives: IV
```{r desc2, echo=FALSE}
iv_colnames <- c("pca1_stock", "med_hh_income_1000", "avg_rooms_per_dwelling", "percent_hh_with_children", "percent_drivers_female", "percent_publictransit_female")
avg_sndi <- mean(df$pca1_stock)
ggplot(csd, aes(pca1_stock)) + geom_histogram(color="black", fill="white") + geom_vline(xintercept=avg_sndi, color='red') + labs(title='Street network disconnectedness in Canada by census tract', x='SNDI') 

st_drop_geometry(df[iv_colnames]) %>% summarise_all(mean) %>% t()
```


## Study 1: Gender differences in spatial behaviour
### Commuting Modes

```{r commute_modes, echo=FALSE }
commute_modes_female <- data.frame(driver=df$commute_driver_female, passenger=df$commute_passenger_female, transit=df$commute_publictransit_female, walk=df$commute_walk_female) 
commute_modes_female$gender <- 'Female'
commute_modes_male <- data.frame(driver=df$commute_driver_male, passenger=df$commute_passenger_male, transit=df$commute_publictransit_male, walk=df$commute_walk_male) 
commute_modes_male$gender <- 'Male'
commute_modes <- rbind(commute_modes_female, commute_modes_male)
ggplot(melt(commute_modes, id.vars='gender'), aes(x=value, y=variable, fill=gender)) + geom_boxplot() + facet_wrap(~variable, scale="free") + coord_flip() + labs(x=NULL, y=NULL, fill='Gender')
```

\[H_0: \text{driver}_{F} \geq \text{driver}_M \\ H_1: \text{driver}_{F} \lt \text{driver}_M\] 

```{r}
t.test(commute_modes_female$driver, commute_modes_male$driver, alternative='less', var.equal=TRUE, paired=TRUE)
```
Reject null hypothesis; less women commute as drivers than men. 

\[H_0: \text{transit}_{F} \leq \text{transit}_M \\ H_1: \text{transit}_{F} \gt \text{transit}_M\]

```{r}
t.test(commute_modes_female$transit, commute_modes_male$transit, alternative='greater', var.equal=TRUE, paired=TRUE)
```
Reject null hypothesis; more women commute by public transit than men. 


### Commuting Durations
```{r commute_time, echo=FALSE }
commute_time_female <- data.frame(t15=df$commute_time_lt15_female, t15to29=df$commute_time_15to29_female, t30to44=df$commute_time_30to44_female, t45to59=df$commute_time_45to59_female, t60=df$commute_time_gt60_female) 
commute_time_female$gender <- 'Female'
commute_time_male <- data.frame(t15=df$commute_time_lt15_male, t15to29=df$commute_time_15to29_male, t30to44=df$commute_time_30to44_male, t45to59=df$commute_time_45to59_male, t60=df$commute_time_gt60_male) 
commute_time_male$gender <- 'Male'
commute_times <- rbind(commute_time_female, commute_time_male)
ggplot(melt(commute_times, id.vars='gender'), aes(x=value, y=variable, fill=gender)) + geom_boxplot() + facet_wrap(~variable, scale="free") + coord_flip() + labs(x=NULL, y=NULL, fill='Gender')
```

T-tests: all stat sig that male and female are diff
```{r commute_t}
t.test(commute_time_female$t15, commute_time_male$t15, var.equal=TRUE, paired=TRUE)
t.test(commute_time_female$t15to29, commute_time_male$t15to29, var.equal=TRUE, paired=TRUE)
t.test(commute_time_female$t30to44, commute_time_male$t30to44, var.equal=TRUE, paired=TRUE)
t.test(commute_time_female$t45to59, commute_time_male$t45to59, var.equal=TRUE, paired=TRUE)
t.test(commute_time_female$t60, commute_time_male$t60, var.equal=TRUE, paired=TRUE)
```


## Study 2: determinants of LFP
```{r lfp_all}
df_vars <- df_no_geom[iv_colnames]
df_vars$lfp_female <- df_no_geom$lfp_female
model_all <- lm(lfp_female ~ ., data=df_vars)
summary(model_all)
plot(residuals(model_all), ylab='residuals')
plot(residuals(model_all) ~ df_vars$lfp_female, ylab='residuals', xlab='fitted value')
qqnorm(residuals(model_all), main='Residual QQ Plot (normal dist)')
qqline(residuals(model_all))
bptest(model_all)  # heteroskedasticity
model_no_sndi <- lm(lfp_female ~ . -pca1_stock, data=df_vars)
anova(model_no_sndi, model_all)
```

### Multilevel model
```{r lfp_multilevel}
df_vars$cma_uid <- df$cma_uid
model_cma <- lmer(lfp_female ~ pca1_stock + med_hh_income_1000 + avg_rooms_per_dwelling + percent_hh_with_children + percent_drivers_female + percent_publictransit_female + (1 | cma_uid), data=df_vars)
summary(model_cma)
plot(residuals(model_cma), ylab='residuals')
plot(residuals(model_cma) ~ df_vars$lfp_female, ylab='residuals', xlab='fitted value')
qqnorm(residuals(model_cma), main='Residual QQ Plot (normal dist)')
qqline(residuals(model_cma))
```

### Gender Difference in LFP
```{r lfp_diff}
df_vars$lfp_gap <- df_vars$lfp_male - df_vars$lfp_female
model_gap <- lm(lfp_gap ~ . -lfp_female -lfp_male -cma_uid, data=df_vars)
summary(model_gap)
plot(residuals(model_gap), ylab='residuals')
plot(residuals(model_gap) ~ df_vars$lfp_female, ylab='residuals', xlab='fitted value')
qqnorm(residuals(model_gap), main='Residual QQ Plot (normal dist)')
qqline(residuals(model_gap))
```

### Montreal
```{r lfp_mtl, echo=FALSE }
plot(mtl_data['lfp_gap'], breaks='quantile')
plot(mtl_data['pca1_stock'], breaks='quantile')
mtl_data_reg <- mtl_data[iv_colnames]
mtl_data_reg <- st_set_geometry(mtl_data_reg, NULL) 
mtl_data_reg$lfp_gap <- mtl_data$lfp_gap
model_mtl <- lm(lfp_gap ~ . -lfp_male, data=mtl_data_reg)
summary(model_mtl)
plot(residuals(model_mtl), ylab='residuals')
plot(residuals(model_mtl) ~ mtl_data_reg$lfp_gap, ylab='residuals', xlab='fitted value')
qqnorm(residuals(model_mtl), main='Residual QQ Plot (normal dist)')
qqline(residuals(model_mtl))
bptest(model_mtl)  # heteroskedasticity

mtl_data$residuals <- model_mtl$residuals
ggplot(mtl_data) + geom_sf(aes(fill = residuals), colour = "transparent") +   scale_fill_viridis_c(limits = c(-15, 20)) 
```

### Toronto CSD
```{r lfp_to, echo=FALSE }
plot(to_data['lfp_gap'], breaks='quantile')
plot(to_data['pca1_stock'], breaks='quantile')
to_data_reg <- to_data[iv_colnames]
to_data_reg <- st_set_geometry(to_data_reg, NULL) 
to_data_reg$lfp_gap <- to_data$lfp_gap
model_to <- lm(lfp_gap ~ . -lfp_male, data=to_data_reg)
summary(model_to)
to_data$residuals <- model_to$residuals
ggplot(to_data) + geom_sf(aes(fill = residuals), colour = "transparent") +   scale_fill_viridis_c(limits = c(-15, 20)) 
```