---
title: "Data Exploration"
author: "Su Goh"
date: "12/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and do simple linear regression
```{r}
df <- read.csv('data/CA16_CTs_all.csv')
model_general <- lm(lfp_female ~ pca1_stock, data=df)
summary(model_general)
plot(df$lfp_female ~ df$pca1_stock)
abline(model_general)
plot(model_general$residuals)
```
Statistically significant at 5% but the plot shows us that it's not a very linear relationship... Residuals seem to be about 0 so we're just going to pretend that it is linear.

## Montreal
```{r}
mtl_data <- df[df$region_name == "Montréal",]
model_mtl <- lm(lfp_female ~ pca1_stock, data=mtl_data)
plot(mtl_data$lfp_female ~ mtl_data$pca1_stock, main="Montreal", xlab="SNDi", ylab="Female LFP")
abline(model_mtl)
summary(model_mtl)
```
SNDI has a pretty strong effect on female LFP. 

## Toronto
```{r}
to_data <- df[df$region_name == "Toronto",]
model_to <- lm(lfp_female ~ pca1_stock, data=to_data)
plot(to_data$lfp_female ~ to_data$pca1_stock, main="Toronto", xlab="SNDi", ylab="Female LFP")
abline(model_to)
summary(model_to)
```

## Vancouver
```{r}
van_data <- df[df$region_name == "Vancouver",]
model_van <- lm(lfp_female ~ pca1_stock, data=van_data)
plot(van_data$lfp_female ~ van_data$pca1_stock, main="Vancouver", xlab="SNDi", ylab="Female LFP")
abline(model_van)
summary(model_van)
```
Even stronger effect on LFP than in Montreal and Toronto!

## Commuting behaviour
### Mode
```{r, echo=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
boxplot(df$commute_car_male, df$commute_car_female, main="Car", names=c("Male", "Female"))
boxplot(df$commute_publictransit_male, df$commute_publictransit_female, main="Public Transit", names=c("Male", "Female"))
```
do t test?

### Duration
```{r, echo=FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
boxplot(df$commute_time_lt15_male, df$commute_time_lt15_female, main="< 15 minutes", names=c("Male", "Female"))
boxplot(df$commute_time_15to29_male, df$commute_time_15to29_female, main = "15 - 29 minutes", names=c("Male", "Female"))
boxplot(df$commute_time_30to44_male, df$commute_time_30to44_female, main = "30 - 44 minutes", names=c("Male", "Female"))
boxplot(df$commute_time_45to59_male, df$commute_time_45to59_female, main = "45 - 59 minutes", names=c("Male", "Female"))
boxplot(df$commute_time_gt60_male, df$commute_time_gt60_female, main = "> 60 minutes", names=c("Male", "Female"))
```

## Big model
```{r}
model_multi <- lm(df$lfp_female ~ df$avg_hh_size 
                  + df$hh_with_children_single_parents 
                  + df$hh_with_children_couples
                  + df$med_hh_income_thousands
                  + df$avg_rooms_per_dwelling
                  + df$commute_car_female 
                  + df$commute_publictransit_female
                  + df$commute_time_lt15_female
                  + df$commute_time_15to29_female
                  + df$commute_time_30to44_female
                  + df$commute_time_45to59_female
                  + df$commute_time_gt60_female
                  + df$pca1_stock)
summary(model_multi)
```

Hmmm...... all stat sig?? something's not right
